Group 9 Paper

### 2021-12-16 paper.Rmd
Adding final edits to O's methods

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(dplyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls) 
library(tree)
library(randomForest)
library(knitr)
library(class)
library(fastDummies)
library(reprtree)
library(MASS)
library(e1071)
library(gbm)

train <- read.csv("train.csv")
test <- read.csv("test.csv")

train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)
train$HeartDisease <- as.factor(train$HeartDisease)

test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)
test$HeartDisease <- as.factor(test$HeartDisease)
```

@@

@@

@@

```{r,echo=FALSE}
model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
summary(model_glm)

predicted2 <- predict(model_glm, newdata = test, "response")
predicted2 <- ifelse(predicted2>0.5,1,0)
confm2<-table(predicted=predicted2,True=test$HeartDisease)
confm2

paste("Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))

overall_logit <- (confm2[1,1]+confm2[2,2])/sum(confm2)
```

 We started by fitting a logistic regression model with all the variables, and from the results above we see that there are a few non-significant variables and important ones, it is worth noting that SexM variable is is statistically significant and positive, which interesting as it means that on average men are more likely to get heart disease than women given that the other variables are the same, also, ST_SlopeFlat and ST_SlopeUp are statistically significant and the  show different signs which means that people with ST_SlopeFlat have a higher Likelihood of having heart Disease when compared to those that have ST_SlopeUp, and this is consistent with science as ST_Slope is a important factor in the diagnosis of a heart Disease. Overall, for a simple method that also allows for inference, it performed relatively well with an accuracy of 0.85.

@@

K-nearest Neighbors is useful for both regression and classification. An object will be classified by a vote of the nearest k neighbors. The object is assigned to the class most common to those k neighbors. For this project, k was selected by cross validation.  The best accuracy comes from a k of 23 with and accuracy of 0.6878024. Technically both k=1 and k=4 have higher accuracy, but we are avoiding k=1 using as any outlines would create a skew and in certain types of classification problems, our research shower it was better to avoid using an odd number as this can tie votes. With an prediction rate of only 0.6757493 had the lowest overall correct prediction rate.

@@

When using the LASSO method, it was important to select family = "binomial" for the classification problem. The tuning lambda value was selected through cross validation.

@@

This will be the section on Bagging. Include discussion on how & why you chose to do it the way you did it.

@@

```{r,echo=FALSE}
train <- read.csv("train.csv")
test <- read.csv("test.csv")

train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)

test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)

set.seed(445)

oob = trainControl(method = "oob")
cv_5 = trainControl(method = "cv", number = 5)

gbm_grid =  expand.grid(interaction.depth = 1:5,
                        n.trees = (1:5) * 200,
                        shrinkage = c(0.001,0.01,0.015,0.1),
                        n.minobsinnode = 10)

gbm_tune = train(as.factor(HeartDisease) ~ ., data = train,
                      method = "gbm",
                      trControl = cv_5,
                      verbose = FALSE,
                      tuneGrid = gbm_grid)
plot(gbm_tune)

gbm_tune$bestTune

In order to tune tree boosting method I used a 5 fold cross-validation on the training data set, from the plot it seems that there is a spike on the on the accuracy when using shrinkage with 0.01, 400 trees and a interaction depth of 2. It should also be noted that several other ranges of the tune parameters were tested but for simplicity sake on the range on the plot was left which included the best one overall.

```{r,echo=FALSE}
boost <- gbm(HeartDisease~.,data=train,n.trees=400,shrinkage=0.01,interaction.depth = 2 ,distribution = "bernoulli")

boostpred = ifelse(predict(boost,test, n.trees = 400, "response") > 0.5, "1", "0")

confboost<-table(predicted = boostpred, actual = test$HeartDisease)
confboost
error<-(confboost[1,1]+confboost[2,2])/sum(confboost)
overall_boost <- error
paste("accuraccy:",overall_boost)
```
Fitting the model with the tuning parameters previously discussed, we get a much better accuracy when compared to the other methods but at the cost of having lost the ability of doing inference in this data set. We can take a look at the importance graph below, in which we can see that ST_Slope, ChestPainType, and Cholesterol are the most important variables, which is somewhat consistent with most of the models used so far. Also Consistent with the other models, we can see that RestingECG and RestingBP are the least important variables for boosting.

```{r,include=FALSE}
summa<-summary(boost)
```

```{r,echo=FALSE}
ggplot(summa)+geom_col(aes(x=var,y=rel.inf,fill=var))+coord_flip()+labs(y="Relative Influence", x= " Variable")+guides(fill="none")
```

@@

```{r}
rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)), importance = TRUE, type= "classification", ntrees=1000)

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

confm2<-table(predicted=rf,True=true)
confm2

paste(" Test Accuracy :",(confm2[1,1]+confm2[2,2])/sum(confm2))

overall_RF <- (confm2[1,1]+confm2[2,2])/sum(confm2)
rf_fit

imp<-data.frame(rf_fit$importance)

imp$variables <- rownames(imp)
rownames(imp) <- 1:nrow(imp)

imp <- imp[order(imp$MeanDecreaseAccuracy),]
```

Next, we fitted a Random Forest, in this case I chose the number of predictors randomly sampled at each iteration to be $m=\sqrt{p}$, as this was what was recommended by the class book. For random forests, the test accuracy relatively good at 0.8664 which is good, but not better than some of te other methods. Looking at the variable importance, we have some similar results to the other tree based methods, as ST_Slope, ChestPainType, and Cholesterol seem to be the most important variables. However, in random forests we got a negative value for the RestingBP, which is somewhat consistent with the other methos as RestingBP also did not seem to be a important variable in dertermining heart disease, however, in this case it means that for random forests is seems that including RestingBP decreases the accuracy of the method.

```{r}
ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())+guides(fill="none")
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Motivation

@@

For this project, we proposed to use the various machine learning methods learned this semester in CSU's DSCI 445 Machine Learning class that can be applied to a classification problem.

@@

This will be the section on Logistic Regression. Include discussion on how & why you chose to do it the way you did it.

@@

The next method we used was K-Nearest Neighbors. K-nearest Neighbors is useful for both regression and classification. An object will be classified by a vote of the nearest k neighbors. The object is assigned to the class most common to those k neighbors. For this project, k was selected by cross validation.  The best accuracy comes from a k of 23 with and accuracy of 0.6878024. Technically both k=1 and k=4 have higher accuracy, but we are avoiding k=1 using as any outlines would create a skew and in certain types of classification problems, our research shower it was better to avoid using an odd number as this can tie votes. With an accuracy rate of only 0.6757493 had the lowest overall accuracy rate.

```{r message=FALSE, warning=FALSE,echo=FALSE}
temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))
drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
temp_train <- temp_train[, !(names(temp_train) %in% drops)]
temp_test <-temp_test[, !(names(temp_test) %in% drops)]
#trainknn <- temp_train[-temp_train$HeartDisease]
#testknn <- test[-test$HeartDisease]
num_k <- seq(1:100)
my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)
confm_knn_1<-table(knn_pred_1,test$HeartDisease)
overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)
overall_knn_1
 #confusion matrix
kable(confm_knn_1)
mean(knn_pred_1 !=test$HeartDisease)

@@

After KNN, we used  the least absolute shrinkage and selection operator(LASSO). When using the LASSO method, it was important to select family = "binomial" for the classification problem. The tuning lambda value was selected through cross validation. LASSO performed relatively well with an accuracy of 0.85.

```{r message=FALSE, warning=FALSE}
train_matrix <- model.matrix(HeartDisease ~ ., data = train)
test_matrix <- model.matrix(HeartDisease ~ ., data = test)
grid <- 10^seq(10, -2, length=100)
lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
#plot(lasso)
cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min
lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)
num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]
# MSE values to use:
paste("Tuning Lambda",best_lambda_lasso)
#lasso_MSE
#paste("Number of zeros",num_zeros)
#Confusion Matrix
predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
confm_lasso
#Over all correct:
overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)
kable(confm_lasso) #confusion matrix
overall_lasso #over all correct
```

@@

The bagging model was preformed with 500 trees and identified the most important variable as being ST_Slope followed by Chest Pain Type. Interestingly, resting ECG and Resting BP both had negative importance. This means if we remove those two features, we would improve model performance. As other methods found resting ECG and Resting BP to be among the least important variable, it would be worth taking the time to remove them and see if it improves the accuracy rates for our data. Bagging performed rather disappointingly with an accuracy of 0.84, the lowest of all methods except KNN.

```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}
train <- read.csv("train.csv")
test <- read.csv("test.csv")
train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)
test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)
bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE) 
#Confusion Matrix
predict_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predict_bag>0.5,1,0)
confm_bag<-table(predicted_bag,test$HeartDisease)
#MSE
bagging_mse <- mean((predict_bag - test$HeartDisease)^2) 
#Over all correct:
overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)
kable(confm_bag) #confusion matrix
overall_bag #over all correct
#Visulizaton of bagging
VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))
#sort variable importance descending
#VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]
#visualize variable importance with horizontal bar plot
#barplot(VI_plot$imp.0,
   #     names.arg=rownames(VI_plot),
    #    horiz=TRUE,
     #   col='steelblue',
      #  xlab='Variable Importance')
ggplot(VI)+geom_col(aes(x=var,y=Overall,fill=var))+coord_flip()+guides(fill="none")+
  labs(x="Variable Name",y="Importance")
```

@@

This will be the section on Boosting. Include discussion on how & why you chose to do it the way you did it.

@@

This will be the section on Random Forest. Include discussion on how & why you chose to do it the way you did it.

@@

This will be the section on Representative Tree. Include discussion on how & why you chose to do it the way you did it.
